---
title: "Control structures"
output: html_document
date: '2023-05-06'
---

Control Structures in R
--------------------------------------
The commands for, if and while can be used to create loops.  Vectors are generally preferred to loops but...

Common structures
* if, else:  testing a condition
* for: execute a loop a fixed number of times
* while: repeate a loop *while* condiction is true

```{r}
x <- 10
if(x > 3) {
y <- 10
} else {
y <- 0
}
y
```
This can be used for selecting particular events. For example, if you have a dataframe with US, German and UK data it would be possible to say country <- "Germany" and if(country == "Germany", plot(da[,3])) with column three containing the German data etc.

```{r}
for(i in 1:10){
  print(i)
}
```
This could be used for running through a list of tests that you want to make.  For example. 

```{r}
x <- c(2, 3, 4, 5, 8, 10, 3, 4)
for(i in c(mean, sd)){
  a <- "i"(x)
print(a)
}
```
And while
```{r}
count <- 0
while(count <10){
  print(count)
  count <- count +1
}
```
Obviously, you have to make sure that the while will come to an end, otherwise you will be in an infinite loop. 
