---
title: "Writing functions"
output: html_document
date: "May 2023"
---
## Introduction

One advantage of writing programmes is that you can write your own functions to do things that you are likely to want to repeat.  Once you have successfully completed a difficult piece of code, turn it into a function so that you do not have to write it again. 

Like everything in R, a function is an object with particular characteristics.  For example, if we wanted to write a function that would find the square of any number that we input, 

```{r} 
square <- function(x) {
  x^2
}
square(2)
square(3)
square(4)
```

Note the structure: 

```
function name <- function(argument 1, argument 2){

function details (R code)
}
```


We could extend this to calculate a range of exponentials. 



```{r}
exponential <- function(x, y){
  x^y
}
exponential(2, 3)
exponential(5,2)
```



## Financial values

We can write a function that will allow us to value a security. 

Starting with \£100 and an interest rate of 5\%, we would expect to have 

$$FV = PV \times (1 + R)^n$$

where FV is the future value, PV is the present value, R is the interest rate or rate of return and n is the number of years.  This would be: 


```{r}
FV <- function(PV = 100, R, n = 1) {
  FV = PV*(1 + R)^n
}
a <- FV(R = 0.05)
a
a <- FV(100, 0.06)
a
a <- FV(R = 0.1, n = 2)
a
```

The standard valuation of a financial asset is the present discounted value.  The formula would be: 

$$ PV = \frac{FV}{(1 + R)^n}$$

where PV is the present value, FV is the future value, R is the rate of return or the discount rate and n is the number of years into the future. 



### Practice 

1. See if you can write the function for the present value.  This is just the present value re-arranged.  

2. Create a vector that has a range of discount rates (from 5\% to 10\%) and calculate the change in the present value for each of these rates. 


## Descriptive statistics function

The following function will calculate descriptive statistics, store them in a data frame and present the statistics in a clear format. 

```{r}
mystats <- function(x){
  x <- x[!is.na(x)]
  n <- length(x)
  m <- mean(x)
  s <- sd(x)
  med <- median(x)
  skew <- sum((x-m)^3/s^3)/n
  kurt <- sum((x-m)^4/s^4)/n - 3
  max <- max(x)
  min <- min(x)
  mydescriptive <- round(data.frame(
    "Number" = n,
    "Mean" = m,
    "Median" = med,
    "Std dev" = s,
    "Skew" = skew,
    "Kurtosis" = kurt,
    "Maximum" = max,
    "Minimum" = min), 2)
  return(mydescriptive)
}
mydata <- rnorm(100)
mystats(mydata)
```


## Returns 

Remember our code to calculate returns.  We could turn this into a function. 

```{r}
myreturns <- function(x){
  returns <- c(x[1:length(x)-1]/x[2:length(x)] -1, NA)
  return(returns)
}
myseries <- 10:1
round(myreturns(myseries), 2)
```

### Practice

See if you can change the function to allow the function to work on data that runs from old to new.   

## Additional help

There are some examples and further information about functions [here](https://www.dataquest.io/blog/write-functions-in-r/)

