---
output: 
  html_document: 
    theme: united
---
Quantmod examples
========================================================
There are a large number of packages that have been produced by other people.  These packages hold additional functions that allow you to easily accomplish a range of tasks.  You must ``install`` the package to your machine.  This will save all the functions in a directory.  You must then ``require`` the package to make the functions available for a particular session. 

We will look at the ``quantmod`` package. You can read all about the package on ``Cran``.  

```{r, message=FALSE, warning=FALSE,cache=TRUE}
#install.packages('quantmod')
require(quantmod)
getSymbols("BAC")
plot(Cl(BAC), col = 'black')
head(BAC)
tail(BAC)
head(Ad(BAC))
?quantmod
?OHLC
```
There are a number of built in plotting functions. 
```{r, fig.hight=4, warning=FALSE, message=FALSE, cache=TRUE}
require(quantmod)
barChart(BAC)
addTA(EMA(Cl(BAC)),on=1,col=6)
```

**Practice**

* Automatically download the data for the Ford Motor Company from Yahoo
* Check that the data is okay and find out the start and end dates
* Create a plot of the data


  The getSymbols function can draw data from a number of different sources. Use the ``src`` argument.  

Get Yahoo stock data from Google and Google stock data from Yahoo. 

```{r, warning=FALSE, message=FALSE, cache=TRUE}
require(quantmod)
getSymbols("YHOO",src="google") # from google finance
head(YHOO)
getSymbols("GOOG",src="yahoo") # from yahoo finance
head(GOOG)
```

###Technical trading
The ``ttr`` package is installed and loaded as part of ``quantmod``

```{r, message=FALSE, warning=FALSE}
require(TTR)
require(quantmod)
barChart(YHOO)
candleChart(GOOG,multi.col=TRUE,theme='white', subset = 'last 4 months')
chartSeries(YHOO)
addMACD()
addBBands()
```

**Practice**

* Check the functions that are available in ``TTR`` package
* Create a 5-day moving average of YHOO and call it ``YHOO5DMA``` 
* Plot the Yahoo 5-day moving average. 


###The Fred database
One of the sources is the Fred database.  Here we get the data and chart it in different ways. 
```{r, cache=TRUE, message=FALSE, warning=FALSE}
getSymbols('CPIAUCNS',src='FRED')
plot(CPIAUCNS)
plot(diff(log(CPIAUCNS)))
plot(diff(log(CPIAUCNS), lag=12), main="CPI Inflation")
```

```{r, cache=TRUE}
getSymbols('DEXJPUS',src="FRED") # FX rates from FRED
  plot(DEXJPUS, col = 'black', main = "USD-JPY")
```

**Practice**

* Go to the St. Louis Fed database
* Find the code for US unemployment data
* Plot the data and beautify the chart