First R
========================================================

All the regular mathematical operations can be carried out. 
```{r, eval=FALSE}
#First R session
4+5
4*5
4/5
4^5
sqrt(5)
log(5)

test<-2*3
test

test=2/3
test
```
R Objects
----------------------------------------------------------
R uses objects. The main objects are:
* numeric
* character
* logical 
* Vector
* Data frame

Character
-----------------------------------------------------
For an example of character
```{r}
a <- "My name"
a
```

Vectors and matrices
-----------------------------------------------------
Strings or numerics can be concatonated or combined to form a vector or matrix
```{r}  
mynumbers <- c(3,5,6,7,9)
mynumbers
mynumbers <- c(1:10)
mynumbers
```
To create a sequence, use the command sequence
```{r}
a <- seq(0.5,2.5, length=100)
a
b <- seq(1,10,1)
b
mynumbers <- 1:12
m <-matrix(mynumbers, nrow=4)
m
```
The plot function will allow you to create graphs with the data that you have. 
```{r, fig.height=4}
x<-seq(-pi,pi,length=10)
y<-sin(x)
plot(x,y)
lines(x,y)
```
If you change the length of the x variable to 100 you will get a smoother line. 
```{r, fig.height=4}
x<-seq(-pi,pi,length=100)
y<-sin(x)
plot(x,y)
lines(x,y)
```
```{r, fig.height=4}
z<-rnorm(100)
hist(z, prob=TRUE)
mu<-mean(z)
sig<-sd(z)
x<-seq(-4,4,length=500)
y<-dnorm(x,mu,sig)
lines(x,y, col='red')
```
